---
title: "biooracle"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Extra R-language tools to supplement [biooracler R package](https://github.com/bio-oracle/biooracler). This package serves up R scripts to create a local data repository.

# Requirements for package

From CRAN

  + [R v4.1+](https://www.r-project.org/)
  + [rlang](https://CRAN.R-project.org/package=rlang)
  + [stars](https://CRAN.R-project.org/package=stars)
  + [sf](https://CRAN.R-project.org/package=sf)
  + [dplyr](https://CRAN.R-project.org/package=dplyr)
  + [tidyr](https://CRAN.R-project.org/package=tidyr)
  

# Installation

```
# install.packages(remotes)
remotes::install_github("BigelowLab/biooracle")
```



# Set up a data directory

You can store the path to your chosen data directory. It will persist between R sessions so you don't have to do it each time.

```{r setup_data_dir}
suppressPackageStartupMessages({
  library(biooracle)
  library(dplyr)
})
set_biooracle_root("~/Library/CloudStorage/Dropbox/data/biooracle")
```

We'll be creating a new local dataset for the Northwest Atlantic (nwa) which has the bounding box `{r} bb = c(xmin = -77, xmax = -42.5, ymin = 36.5, ymax = 56.7)`.

```{r make_nwa}
nwa_path = biooracle_path("nwa") |> make_path()
biooracle_path() |> dir(full.names = TRUE)
```

# List available data layers

[biooracler R package](https://github.com/bio-oracle/biooracler) provides a nice utility for obtaining a list of available data layers.  We have enhanced that by parsing the `dataset_id` so we can quickly filter for the desried dataset(s).  This will only work if [biooracler](https://github.com/bio-oracle/biooracler) is installed.

```{r list_layers}
list_layers() |>
  knitr::kable()
```

# Fetch some data to a temporary directory

We'll set that aside for right now and fetch some data for that region, but note that this downloaded as a NetCDF file in a temporary directory.  Keep in mind that we are specifying the
bounding box with a vector of the corners, but we can also provide any object from which a bounding box can be determined using the [sf package](https://CRAN.R-project.org/package=sf), such as a polygon, raster or collection of points.

```{r download, message = FALSE, warning = FALSE}
dataset_id = "thetao_ssp119_2020_2100_depthmin"
newfile = fetch_biooracle(dataset_id, 
                          bb = c(xmin = -77, xmax = -42.5, ymin = 36.5, ymax = 56.7))
```

**NOTE** that you can make subselections of variable and times to download.  See `?fetch_biooracle` for the details.


Now we can read the file.

```{r read}
x = stars::read_stars(newfile, quiet = TRUE)
x
```

# Archiving in a local database

We often save the data in a directory structure aong with a simple table that catalogs the contents of the directory.  The `archive_biooracle()` function will split up the fecthed data and save in a logical data structure.  We provide the data path, in this case for the Northwest Atlantic (nwa).

```{r archive}
archive_biooracle(newfile, path = nwa_path)
```

Alternatively, it is possible to fetch and archive in one step, and this is likely the most convenient usage.

```{r}
newfile = fetch_biooracle(dataset_id, 
                          bb = c(xmin = -77, xmax = -42.5, ymin = 36.5, ymax = 56.7),
                          archive = TRUE,
                          data_dir = nwa_path)
```

# Read the database catalog

Once you have established a database of files, your can read the database catalog.

```{r read_database}
db = read_database(nwa_path) |>
  print()
```


# Read in data from the database

You can use a portion of the database to read in a `stars` object.  Keep in mind that if you are reading multiple over multiple decades, then each variable must have the same number of time steps.

```{r read_biooracle}
x = db |>
  dplyr::mutate(year = as.numeric(year)) |>
  dplyr::filter(year >= 2070) |>
  read_biooracle(, path = nwa_path) |>
  print()
```

And of course you can plot.

```{r plot_again}
plot(x['thetao_mean'])
```
